### 2013-09-12 legend.tex
Merge branch 'master' of svn.ais.uni-bonn.de:u/amueller/icra_2014_nyu
git-svn-id: https://svn.ais.uni-bonn.de/svn/staff/publications/icra_2014_mueller_nyu@4357 b2511e17-9ac0-488b-bc21-1d6d82e5baa2

\documentclass{standalone}
\usepackage{color}
\usepackage{tikz}

\definecolor{structure}{RGB}{185, 202, 202}
\definecolor{prop}{RGB}{227, 172, 0}
\definecolor{furniture}{RGB}{255, 94, 28}
\definecolor{ground}{RGB}{9, 68, 83}
%[``structure'', ``prop'', ``furniture'', ``ground'' , ``void'']

\begin{document}
    \begin{tikzpicture}
        \matrix [column sep=.5cm]{%
        \node [rectangle, fill=ground, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Ground};&
        \node [rectangle, fill=structure, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Structure};&
        \node [rectangle, fill=furniture, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Furniture};&
        \node [rectangle, fill=prop, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Props};&
        \node [rectangle, fill=black, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Void};\\
    };
    \end{tikzpicture}
\end{document}