### 2013-09-12 paper.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\makeatletter
\let\NAT@parse\undefined
\makeatother

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphicx} % for pdf, bitmapped graphics files
\usepackage{epsfig}
\usepackage{refstyle}
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{booktabs}
\usepackage{tabu}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{matrix}

%\usepackage[numbers]{natbib}
\usepackage[natbib, style=ieee]{biblatex}
\addbibresource{paper.bib}

\DeclareMathOperator*{\argmax}{arg\,max}

\title{\LARGE \bf
Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation of RGB-D Images
}

\author{Andreas C. M\"uller and Sven Behnke
\thanks{All authors are with the Autonomous Intelligent Systems Group, Univer-
        sity of Bonn, Germany. Email: amueller@ais.uni-bonn.de
    }}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
    We present a structured learning approach to semantic annotation of RGB-D
    images. 
    Our method learns to reason about spacial relations of objects and
    fuses low-level class predictions to a consistent interpretation of a
    scene. Our model incooporates color, depth and 3D scene features, on which
    an energy function is learned to directly optimize object class prediction using
    the loss-based maximum-margin principle of structural support vector machines.
    %With the proliferation of cheap % TODO urghs formulation
    %high-quality RGB-D sensors, it became possible to capture more and more
    %data not only as color images, but with additional depth information. The
    %depth channel provides important clues not only for scene structure, but
    %also for predicting semantic object classes.
    %We demonstrate how conditional random field learning, an approach
    %that has shown great promise for RGB images, can exploit depth information
    %to improve semantic interpretation of the scene.
    We evaluate our approach on the NYU V2 dataset of indoor scenes, a challenging
    dataset covering a wide variety of scene layouts and object classes. We
    hard-code much less information about the scene layout into our model then previous
    approaches, and instead learn object relations directly from the data. We
    find that our conditional random field approach improves upon previous
    work, setting a new state-of-the-art for the dataset.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}
For robots to perform varied tasks in unstructured environments, understanding their
sourroundings is essential. We formulate the problem of semantic annotation of maps
as a dense labeling of RGB-D images into semantic classes. Dense labeling of objects
and structure classes allows for a detailed reasoning about the scene.

In this work, we propose the use of random forests combined with conditional
random fields (CRFs) to perform robust estimation of structure classes in RGB-D
images. The CRF is learned using a structural support vector machine,
allowing it to integrate the noisy low-level information produced by a pixel-based
random forest to a consistent interpretation of the scene.

We thereby extend the success of learned conditional random field models for
semantic segmentation in RGB images to the domain of 3D scenes.
Our emphasis lies on expoiting the additional depth and 3D information in all
procsessing steps, while relying on \emph{learning} to create a model that 
is adjusted to the properties of the sensor input and environment.

Our approach starts with a random forest, providing a noisy local estimate
of semantic classes based on color and depth information. These estimates
are grouped together using a superpixel approach, for which we extend previous
superpixel algorithms from the RGB to the RGB-D domain.
We then build a geometric model of the scene, based on the neighborhood graph
of superpixels.  We use this graph not only to capture spacial relations in the
2d plane of the image, but also to model object distances and surface angles in
3d, using a point cloud generated from the RGB-D image.

We access the accuracy of our model on the challenging NYU Segmentation V2
dataset \citep{SilbermanECCV12}, where our model outperforms previous
approaches.  Our analysis shows that while our random forest model
already has competative performance, the superpixel-based grouping and in
pariculare the loss-based learning are integral ingredients of the success of our method.

\section{RELATED WORK}
The task of dense semantic labeling in natural images has been widely studied
in computer vision.  Our approach of using pixel-based low-level cues is
similar to the work of \citet{shotton2006textonboost}, who introduced
TextonBoost for semantic segmentation, and used a hand-coded CRF energy to
model spacial interactions. More recent work usually employ learned conditional
random fields, learning object interactions from the
data to directly maximize a loss function~\citep{lucchi2013learning, krahenbuhl2012efficient},
which is the approach we are following.

Work using RGB-D data only recently emmerged with
the availability of the NYU Depth Dataset V1~\citep{silberman2011indoor}.
\citet{silberman2011indoor} captured a wide variety of indoor scence using a structured light sensor,
and provided dense semantic object class annotation. Their work also introduced a baseline method
for semantic segmentation of RGB-D image, which used an approach based on superpixels, interest
point descriptors and a conditional random field. While pairwise potentials for the CRF were carfully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our work, which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
\citet{ren2012rgb} focus on the design of features for semantic labeling of RGB-D data, and use
a hierarchical segmentation to provide context. While they also define a conditional random field on superpixels,
their model is again not learned, but a weighted potts model, using only a
probability of boundary map, and not taking spacial layout into account at all.

\citet{SilbermanECCV12} extendet the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using in this work.  Their focus is on inferring support relations in indoor
scence, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} use a complex pipeline, employing significant domain knowledge.
In this work, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

\citet{couprie-iclr-13} approach the task of semantic segmentation of structure classes
in RGB-D using the paradigm of convolutional neural networks, extending
previous work of \citet{farabet-pami-13} and \citet{schulz2012learning}.
Similar to our approach, \citet{couprie-iclr-13} combine the output of a
pixel-based, low-level learning algorithm with an independent unsupervised
segmentation step. In contrast to their work, we improve our results by not
only averaging predictions within superpixels, but also explicitly learning
interactions between neighboring superpixels, favouring a consistent
interpretation of the whole image.

\citet{stueckler2013} proposed a real-time system for online learning and prediction of semantic classes.
Their main contribution is the online integration of low-level predictions over multiple frames
by estimating camera motion and building a 3D model of the scene.
We use the implementation of random forest provided by \citet{stueckler2013}, but instead of
integrating predictions over time, we focus on exploiting the structur within a single frame.

\section{LEARNING DEPTH-SENSITVE CONDITIONAL RANDOM FIELDS}
We take a conditional random field approach, whose nodes repesent a labeling of
superpixels.  We use an energy consisting of first and second order factors
(also called unary and pairwise potentials), with learned potential functions.
Let us denote the representation of an input image by $x$ and a labeling of
superpixels into semantic classes by $y$.  Then the general form of the energy is
\begin{equation}\eqlabel{energy}
    g(x, y) = \sum_{v \in V} \psi_v(x, y_v) + \sum_{(v, w) \in E} \psi_{v,w}(x, y_{v}, y_{w}).
\end{equation}
Here $V$ enumerates the superpixels, and $E\subset V \times V$ is a set of
edges, encoding neighborhood between superpixels.

We learn the unary and pairwise energy functions $\psi_v$ and $\psi_{v, w}$ from the training data
using a structural support vector machine. 
The concept of structural support vector machines allows for a principled,
maximum-margin based, loss-sensitive training of conditional random fields, that allows much
more complex interactions than the simple potts potentials that are often used in the literature.

In general, structural support vector machines learn the parameters $\theta$ of a predictor of the form
\begin{equation}\eqlabel{prediction}
f(x) = \argmax_{y \in \mathcal{Y}} \theta^T \Phi(x, y).
\end{equation}
We choose $\psi$ in \Eqref{energy} to be linear in the learnable parameters and
the data dependend features, resulting in a form equivalent to
\Eqref{prediction}. Our features are described in detail below.  We use the
$1$-lack formulation of the structural SVM~\citep{joachims2009cutting} and
solve the maximization in \Eqref{prediction} using a combination of fusion
moves~\citep{lempitsky2010fusion} and the AD$^3$ algorithm of
\citet{martins2011augmented}.  In contrast to graph-cut inference, fusion moves
and AD$^3$ can work with arbitrary potential function, and allow percise
learning using the structural support vector machine approach.

\subsection{Low Level Segmentation}
We take a super-pixel based approach to semantic segmentation.
Our superpixel generation is based on the SLIC algorithm~\citep{achanta2012slic}. We extend the
standard SLIC algorithm, which works on the Lab space to also include 
depth information. The resulting algorithm is a localized $k$-Means in Lab-D-XY space.
Our implementation is publicly available through the scikit-image library\footnote{\url{http://scikit-image.org}}.
Similar to \citet{SilbermanECCV12}, we found little visual improvement over the
RGB segmentation when using additional depth information. On the other hand,
estimation of per-superpixel features based on the 3D point cloud was more
robust when including depth information into the superpixel procedure.
%TODO visualize?
The resulting superpixels are compact in the 2D image. As the corresponding point
cloud is very sparse at depth jumps, it is unfortunately not possible to create
superpixels that are compact in 3D while maintaining a meaningful minimum size.

\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/height_success}\\
        \vspace{3mm}
        \includegraphics[width=\linewidth]{images/height_failure}
    \end{center}
    \caption{%
        Visualization of the height computed using the method described in \Secref{unaries}.
        Input images are shown on the left (depth not shown), the computed height is depicted on the right.
        The top row exemplifies a typical scene, while the bottom row shows a scene without horizontal
        surfaces, where our method fails.
    \figlabel{height}}
\end{figure}

\subsection{Unary Image Features}\seclabel{unaries}
Our method builds on the probability output of a random forest, trained for pixelwise
classification of the structure classes.
We use the GPU implementation provided by \citet{stueckler2013}.
The input for training are the full RGB-D images, transformed to Lab
color space. Each tree in the forest uses training pixel only from a subset of
training images. For each training image, an equal number of pixels for each
occurring class is sampled. Split features are given by difference of regions
on color or depth channels. Region size and offsets are normalized using the
depth at the target pixel.
We accumulate the probabilistic output for all pixels within a superpixel, and
use the resulting distribution as a feature for the unary node potentials in our CRF model.
We augment these prediction with another feature, based on the height of a superpixel in 3D.
This is a very informative feature, in particular to determin the floor.
To compute the height of a (super) pixel, we first find the ``up'' direction.
We use a very simple approach that we found effective: we cluster normal directions of all pixels
into 10 clusters using $k$-means, and use the one that is most parallel to the $Y$ direction, which
roughly corresponds to height in the dataset.
We then project the 3D pointcloud given by the depth along this direction, and normalize the result between
$0$ and $1$.
This procedure works robustly given there is any horizontal area in the image,
such as the ground or a table. A few scenes contain only walls and furniture,
and the approach fails for these. \Figref{height} illustrates a typical case and
one of the much rarer failure cases.
While we could use a more elaborat scheme,
such as the one from \citet{SilbermanECCV12}, we suspect that the feature is of
little use in scenes without horizontale surfaces.
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/normal_feature}
    \end{center}
    \caption{%
        Visualization of one of the pairwise features, the similarity between superpixel normals.
        The image shows the zoom-in of a bedroom scene, together with the superpixel oversegmentation.
        Lines connect adjacent superpixels, and line-stregth gives the magnitude of the orientation similarity.
    \figlabel{normal_feature}}
\end{figure}

\subsection{Pairwise Depth-Sensitive Features}
There are five different features used to build pairwise potentials in our model:
\begin{description}
    \item[Constant] \hfill \\
        A constant feature allows to model general neighborhood relations.

    \item[Color Contrast] \hfill \\
        We employ a non-linear color contrast, as is common in the computer
        vision literature, between the superpixel mean colors $c_i$ and $c_j$:
        \[\exp\left(-\gamma \lVert c_i - c_j \rVert^2\right)\]

    \item[Vertical Alignment] \hfill \\
        We model the directed angle between superpixel centers in the 2d image
        plane.  This allows the  model to learn that ``structure'' is above
        ``floor'', but not the other way around.

    \item[Depth Difference] \hfill \\
        We include the signed depth difference between superpixels, which
        allows the model to detect depth discontinuities that are not
        represented in the 2d neighborhood graph of the superpixels.

    \item[Normal Orientations] \hfill \\
        Differences in normal vector orientation are a strong clue on
        whether two superpixels belong to the same surface, and therefore the
        same structural class.
        We compute the orientation of normals in the 3D using the method of \citet{holz_2011_robocup},
        as implemented in the point cloud library (pcl)\footnote{\url{http://pointclouds.org}}.
        All normals within a superpixel are then averaged, to get a single orientation for each superpixel.
        The feature is computed as $\frac{\pi}{4}$ minus the (undirected) angle between the normals belonging
        to two adjacent superpixels.  
        The feature is illustrated in \Figref{normal_feature}. The difference
        in normal orientation highlights that pillow and wall are distinct
        objects, even though there is no strong distinction in color or depth.
\end{description}

\section{EXPERIMENTS}

\begin{table*}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
\toprule
                        & ground        &  structure    & furniture     & props         & class average   & pixel average\\
\cmidrule(r){1-7}

RF                      &         90.8  &   81.6        & 67.9          & 19.9          &  65.0        &  68.3 \\
RF + SP                 &         92.5  &   83.3        & 73.8          & 13.9          &  65.7        &  70.1 \\ 
RF + SP + SVM           &         94.3  &   78.3        & 62.5          & \textbf{46.0} &  70.3        &  69.8 \\
RF + SP + CRF           &         94.8  &   78.7        &          70.8 & 41.6          &\textbf{71.5} &  \textbf{72.0} \\
\cmidrule(r){1-7}
\citet{SilbermanECCV12} &         68    &   59          & 70           & 42            &  59.6        & 58.6 \\
\citet{couprie-iclr-13} &         87.3  & \textbf{86.1} & 45.3         & 35.5          &  63.5        & 64.5 \\
\citet{stueckler2013}   & \textbf{95.6} &   83.0        & \textbf{75.1}& 14.2          &  67.0        & 70.9 \\

\bottomrule
\end{tabularx}
    \caption{Quantitative comparison of the proposed method with the
literature. The best value in each column is printed in bold. The upper part of
the table shows contributions by different parts of our pipeline. RF stands for random forest prediction, RF + SP for aggregated
random forests prediction within superpixels, RF + SP + SVM for an SVM trained on the unary potentials, and RF + SP + CRF is
our proposed pipeline. We optimized our approach for class average
accuracy.\tablabel{results}}
\end{table*}

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 indoor
RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), which puts more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor.

Our approach is implemented using our \textsc{PyStruct}
library~\footnote{\url{http://pystruct.github.io}}.  The code for our
experiments will be released upon acceptance. We use \textsc{scikit-learn}~\citep{pedregosa2011scikit}
for $k$-means clustering and the SVM baseline.

All hyper parameters were adjusted using 5-fold cross-validation. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework of \cite{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using standard grid-search.
We found 500 superpixels per image to work best, which allow for a maximum
performance of 95\% average class accuracy on the validation set.

\subsection{Results}
\Tabref{results} compares different components of our approach with the literature.
Please note that we \emph{first} designed our final model, using only the
validation data. We now report accuracies of simpler models for reference,
however these results were not used for model selection. To separate
the influence of loss-based training and the spacial reasoning of the conditional random field,
we also train a standard support vector machine on the unary potentials for comparison.

The random forest prediction, as reported in \citet{stueckler2013} is already quite competitive.
Grouping into superpixels slightly improves performance, by removing
high-frequency noise and snapping to object boundaries. Somewhat surprisingly,
using a standard unstructured SVM with rescaled loss already advances the mean
accuracy decidedly above the previous state-of-the art. We
attribute this mostly to the ability of the SVM to more directly minimize the
target loss, leading to a calibration for the actual error measure. On the
other hand, the total accuracy only decreases by $0.3\%$, while the average
accuracy increases by $4.5\%$. This means that the SVM can further exploit
uncertainty within superpixels and correlations between classes.
This performance is still improved upon, both in class average and pixel average performance
by the learned conditional random field approach, yielding the best published result
so far for both measures. The increase over the standard SVM is $1.2\%$ for class average accuracy
and $2.2\%$ for pixel average accuracy.

A visualization of the impact of each processing step can be found in
\Figref{comparison}, which shows prediction results on the test set.  The for
prediction methods correspond to the rows of \Tabref{results}.  The difference
between the SVM and CRF approaches are clearly visible, with the CRF producing
results that are very close to the ground truth in serveral complex scenes.

We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the conditional random field can.
On the other hand, performance of the CRF deteriorates sligthly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the conditional random field has little chance to recover.

\begin{figure}
    \begin{center}
        \includegraphics[width=.45\linewidth]{images/vertical_alignment}\vspace{1mm}
    \includegraphics[width=.45\linewidth]{images/depth_difference}
    \end{center}
    \caption{%
        Visualization of some of the learned potentials. The right potential is
        applied to relative depth between superpixels, the second on the
        feature encoding whether one superpixel is above the other in the
        image. See text for details.
    \figlabel{weights}}
\end{figure}

\Figref{weights} visualizes some learned potential functions.
Higher values correspond to favored configurations. One can see that the
vertical alignment potential expresses that the floor is much more likely to be below
other classes. It also encodes the fact that props
rest on furniture, but not the other way around.
The potential of the depth feature encodes, for example, that the ground
is usually behind the other classes, while furniture is in front of structures,
such as the wall.
Interestingly, the potential functions are not anti-symmetric, and forcing them to be so
degrates the results. This suggests that the direction of connecting edges, going from
the top left to the bottom right in the image, is also exploited by the potentials.

\begin{figure*}
    \begin{tabu} to \linewidth{@{}XXXXX@{}}
    \includegraphics[width=\linewidth]{images/00845_image.png}&%
    \includegraphics[width=\linewidth]{images/00781_image.png}&%
    \includegraphics[width=\linewidth]{images/01331_image.png}&%
    \includegraphics[width=\linewidth]{images/00118_image.png}&%
    \includegraphics[width=\linewidth]{images/01203_image.png}\\
    \multicolumn{5}{@{}c@{}}{Input Image}\\

    \includegraphics[width=\linewidth]{images/00845_pixel.png}&%
    \includegraphics[width=\linewidth]{images/00781_pixel.png}&%
    \includegraphics[width=\linewidth]{images/01331_pixel.png}&%
    \includegraphics[width=\linewidth]{images/00118_pixel.png}&%
    \includegraphics[width=\linewidth]{images/01203_pixel.png}\\
    \multicolumn{5}{@{}c@{}}{Random Forest Prediction}\\

    \includegraphics[width=\linewidth]{images/00845_sp.png}&%
    \includegraphics[width=\linewidth]{images/00781_sp.png}&%
    \includegraphics[width=\linewidth]{images/01331_sp.png}&%
    \includegraphics[width=\linewidth]{images/00118_sp.png}&%
    \includegraphics[width=\linewidth]{images/01203_sp.png}\\
    \multicolumn{5}{@{}c@{}}{Superpixel Voting}\\

    \includegraphics[width=\linewidth]{images/00845_svm.png}&%
    \includegraphics[width=\linewidth]{images/00781_svm.png}&%
    \includegraphics[width=\linewidth]{images/01331_svm.png}&%
    \includegraphics[width=\linewidth]{images/00118_svm.png}&%
    \includegraphics[width=\linewidth]{images/01203_svm.png}\\
    \multicolumn{5}{@{}c@{}}{Support Vector Machine on Superpixels}\\

    \includegraphics[width=\linewidth]{images/00845_ssvm.png}&%
    \includegraphics[width=\linewidth]{images/00781_ssvm.png}&%
    \includegraphics[width=\linewidth]{images/01331_ssvm.png}&%
    \includegraphics[width=\linewidth]{images/00118_ssvm.png}&%
    \includegraphics[width=\linewidth]{images/01203_ssvm.png}\\
    \multicolumn{5}{@{}c@{}}{Conditional Random Field on Superpixels}\\

    \includegraphics[width=\linewidth]{images/00845_gt.png}&%
    \includegraphics[width=\linewidth]{images/00781_gt.png}&%
    \includegraphics[width=\linewidth]{images/01331_gt.png}&%
    \includegraphics[width=\linewidth]{images/00118_gt.png}&%
    \includegraphics[width=\linewidth]{images/01203_gt.png}\\
    \multicolumn{5}{@{}c@{}}{Ground Truth}\\
    \multicolumn{5}{@{}c@{}}{
        \includegraphics[]{legend.pdf}
    }
    \end{tabu}
\caption{%
Qualitative evaluation of the conditional random field.
The first three images illustrate errors in the original prediction that can be corrected, while
the second two images illustrate failure modes. See the text for details. Pixels marked as void
are excluded from the evaluation.
\figlabel{comparison}}
\end{figure*}

\section{SUMMARY AND DISCUSSION}
We introduce a conditional random field formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spacial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function that is fully learned using 
 the structural support vector machine approach.  By explicitly modeling 3D
 relations in a fully learned framework, we improve the state-of-the art on the
 NYU V2 dataset for semantic annotation of structure classes.

While our approach allows modelling of spacial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models~\citep{dann2012pottics} or higher
order potentials~\citep{kohli2009robust}.
Another possible line of future work is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model, as done by
\citet{SilbermanECCV12}.

While this work did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citep{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.

\printbibliography{}

\end{document}